Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data

### 2020-06-12 project.Rmd
Last fixes before delivery

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + hours(blocks$start + 1)
blocks$end <- seis.first.second + hours(blocks$end + 1)
plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start (mm-'yy)",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
blocks$start
blocks$interval
remove(seis.complete, seis.missing.ids, seis.present, seis.total, seis.missing, block_start, block_end, blocks)

@@

## Periodgram
```{r periodgram}
min_h_seasonality <- 24
min_h_seasonality <- 2
pgram <- spec.pgram(stats::filter(build_ts(seis.h, start = seis.train.start, end = seis.train.end), side=2, filter=c(0.5, rep(1,min_h_seasonality-1), 0.5)/min_h_seasonality),
#`pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end),
                    detrend=F, demean=T, na.action=na.aggregate, plot=F)
pgram.data <- data.frame(freq=pgram$freq, spec=pgram$spec)
ggplot(pgram.data) +
  geom_line(aes(freq, spec)) +
  xlab("frequency") + ylab("spectrum") +
  labs(title = "Periodgram", caption = paste0("bandwidth = ", pgram$bandwidth, "    df = ", pgram$df)) ->
  periodgram

start_ratio <- 2
filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*start_ratio)) 
top95pc <- pgram$spec[filter_freq]
top95freq <- pgram$freq[filter_freq][order(top95pc, decreasing = T)]
top95pc <- pgram$spec[filter_freq][order(top95pc, decreasing = T)]
cat("SD ratio\n")
(top95pc - mean(pgram$spec))/sd(pgram$spec)
top95seas.s <- 1/top95freq
cat("Raw frequency by unit\n")
top95seas.s
top95seas.h <- top95seas.s/3600
cat("Frequency by unit\n")
top95seas.h

ratio.color <- function(x) ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
shown_labels <- 4
periodgram +
  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 3, 2*c(1,2,3,4,6,12), 24*7), labels = as.integer, name="hours")) +
  scale_y_continuous(sec.axis = sec_axis(~., breaks=mean(pgram$spec) + sd(pgram$spec) * c(1,3), labels=c("1", "2", "3")[c(1,3)], name="spectrum/sd ratio")) +
  geom_hline(yintercept=(mean(pgram$spec) + sd(pgram$spec) * c(1:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(1:3))) +
  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, (spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:shown_labels,], aes(freq, (spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
  scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])

>>>>

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + hours(blocks$start + 1)
blocks$end <- seis.first.second + hours(blocks$end + 1)
plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start (mm-'yy)",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval) - .2, labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
# text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$end, y=hour(blocks$interval) - .2, labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=4)
#text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
#text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
blocks$start
blocks$interval
remove(seis.complete, seis.missing.ids, seis.present, seis.total, seis.missing, block_start, block_end, blocks)

@@

## Periodgram
```{r periodgram}
min_h_seasonality <- 24
min_h_seasonality <- 2
pgram <- spec.pgram(stats::filter(build_ts(seis.h, start = seis.train.start, end = seis.train.end), side=2, filter=c(0.5, rep(1,min_h_seasonality-1), 0.5)/min_h_seasonality),
#`pgram <- spec.pgram(build_ts(seis.h, start = seis.train.start, end = seis.train.end),
                    detrend=F, demean=T, na.action=na.aggregate, plot=F)
pgram.data <- data.frame(freq=pgram$freq, spec=pgram$spec)
ggplot(pgram.data) +
  geom_line(aes(freq, spec)) +
  xlab("frequency") + ylab("power level") +
  labs(title = paste("Periodgram -", description), caption = paste0("bandwidth = ", pgram$bandwidth, "    df = ", pgram$df)) ->
  periodgram

start_ratio <- 3
filter_freq <- (pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*start_ratio)) & 1/pgram$freq/3600 < 24*365
top95pc <- pgram$spec[filter_freq]
top95freq <- pgram$freq[filter_freq][order(top95pc, decreasing = T)]
top95pc <- pgram$spec[filter_freq][order(top95pc, decreasing = T)]
cat("SD ratio\n")
(top95pc - mean(pgram$spec))/sd(pgram$spec)
top95seas.s <- 1/top95freq
cat("Raw frequency by unit\n")
top95seas.s
top95seas.h <- top95seas.s/3600
cat("Frequency by unit\n")
top95seas.h

ratio.color <- function(x) ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(x >= (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
shown_labels <- 4
periodgram +
  scale_x_continuous(sec.axis = sec_axis(~1/./3600, breaks=c(1, 3, 2*c(1,2,3,4,6,12), 24*7), labels = as.integer, name="hours")) +
  scale_y_continuous(n.breaks=6) +
  #scale_y_continuous(sec.axis = sec_axis(~., breaks=mean(pgram$spec) + sd(pgram$spec) * c(start_ratio:3), labels=c("1", "2", "3")[c(start_ratio:3)])) +
  geom_hline(yintercept=(mean(pgram$spec) + sd(pgram$spec) * c(start_ratio:3)), linetype="dotted", color = ratio.color(mean(pgram$spec) + sd(pgram$spec) * c(start_ratio:3))) +
  geom_point(data=data.frame(freq=top95freq, spec=top95pc), aes(freq, (spec), color=ratio.color(spec)), shape="o", size=4, show.legend = F) +
  geom_label_repel(data=data.frame(freq=top95freq, spec=top95pc)[1:shown_labels,], aes(freq, (spec), label=round(1/freq/3600), color=ratio.color(spec)), segment.color="grey", show.legend=F) +
  scale_color_manual(values = c("darkgreen", "blue", "red")[start_ratio:3])
